---
title: "Causal Inference Introduction2: Propensity Score Matching"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(weights)
library(MatchIt)
library(knitr)
```

## Load the dataset

Datasets

- Paper: Dehejia R H, Wahba S. Causal effects in nonexperimental studies: Reevaluating the evaluation of training programs[J]. Journal of the American statistical Association, 1999, 94(448): 1053-1062. (http://www.uh.edu/~adkugler/Dehejia&Wahba_JASA.pdf)
- Download: http://users.nber.org/~rdehejia/data/nswdata2.html


```{r load}
# Load the datsets
col.names=c('treat', 'age', 'educ', 'black', 'hispan', 'married', 'nodegree', 're74', 're75', 're78')
dir <- './dataset/nswdata/'
nsw_treated <- fread(paste0(dir, 'nswre74_treated.txt'), col.names = col.names)
nsw_control <- fread(paste0(dir, 'nswre74_control.txt'), col.names = col.names)
cps1_control <- fread(paste0(dir, 'cps3_controls.txt'), col.names = col.names)
cps3_control <- fread(paste0(dir, 'cps3_controls.txt'), col.names = col.names)

# Combine all the datasets
nsw_data_exp <- rbind(nsw_treated, nsw_control)
nsw_data_exp[, dataset := 'NSW-Data-Exp']
nsw_data_obs <- rbind(nsw_treated, cps3_control)
nsw_data_obs[, dataset := 'NSW-Data-Obs']
data <- rbind(nsw_data_exp, nsw_data_obs)
```

Check the sample mean.

```{r pre_treatment}
get.mean.se <- function(x) {
  mean = round(mean(x), 2)
  se = round(sd(x)/sqrt(length(x)), 2)
  return(paste0(mean, '(', se, ')'))
}
results <-  merge(
  data[, .(`no. obs`=.N), by = .(dataset, treat)],
  data[, lapply(.SD, get.mean.se), by = .(dataset, treat)]
)
tmp <- t(results[, 2:ncol(results)])
colnames(tmp) <- results$dataset
kable(tmp)
```

Estimate the Average Treatment Effect on the Treated (ATT).

```{r}
t.test(nsw_data_exp[treat==1, re78], nsw_data_exp[treat==0, re78])
```

## Observational Study: Propensity Score Matching

### Propensity score matching

```{r load}
set.seed(42)
m.out <- matchit(data = nsw_data_obs,
                 formula = treat ~ age + I(age^2) + I(age^3) + educ + 
                   black + hispan + married +
                   I(re74/1000) + I(re75/1000),
                 method = "nearest",
                 distance = "logit",
                 replace = FALSE,
                 caliper = 0.05)
```

### Access Balance

```{r}
plot(m.out, type = "hist", interactive = F)
plot(m.out, type = "QQ", interactive = F, which.xs = c("age", "I(re74/1000)", "I(re75/1000)"))
summary(m.out, standardize = T)$sum.matched
```

### Causal Effect Estimation

```{r}
m.data <- match.data(m.out)

# Direct compare
res <- wtd.t.test(m.data$re78[m.data$treat == 1],
                  m.data$re78[m.data$treat == 0],
                  weight = m.data$weights[m.data$treat == 1],
                  weighty = m.data$weights[m.data$treat == 0])
print(res)
mu <- res$additional[1]
std <- res$additional[4]
cat("Confidence interval: ", sapply(qt(c(0.025, 0.975), coef(res)["df"]), function(x){return(mu+x*std)}), "\n")

# Fit
att.fml <- re78 ~ treat + age + educ + black + hispan + married + nodegree + re74 + re75
fit <- lm(att.fml, data = m.data, weights = m.data$weights)
summary(fit)
cat("Confidence interval: ", confint(fit, "treat", 0.95), "\n")
```


