---
title: 'The NSW Data Playground: Regression'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(MatchIt)
library(knitr)
library(tableone)
```

## 读数据

数据集

- Paper: Dehejia R H, Wahba S. Causal effects in nonexperimental studies: Reevaluating the evaluation of training programs[J]. Journal of the American statistical Association, 1999, 94(448): 1053-1062. (http://www.uh.edu/~adkugler/Dehejia&Wahba_JASA.pdf)
- 下载地址：http://users.nber.org/~rdehejia/data/nswdata2.html

```{r load}
# Load the datsets
col.names=c('treat', 'age', 'educ', 'black', 'hispan', 'married', 'nodegree', 're74', 're75', 're78')
nsw_treated <- fread('./dataset/nswdata/nswre74_treated.txt', col.names = col.names)
nsw_control <- fread('./dataset/nswdata/nswre74_control.txt', col.names = col.names)
cps1_control <- fread('./dataset/nswdata/cps_controls.txt', col.names = col.names)
cps3_control <- fread('./dataset/nswdata/cps3_controls.txt', col.names = col.names)

# Combine all the Datasets
nsw <- rbind(nsw_treated, nsw_control)
cps1 <- rbind(nsw_treated, cps1_control)
cps3 <- rbind(nsw_treated, cps3_control)
nsw[, Dataset := 'NSW']
cps1[, Dataset := 'CPS-1']
cps3[, Dataset := 'CPS-3']
data <- rbind(nsw, cps1, cps3)
data[, Dataset := factor(Dataset, unique(data$Dataset))]
```

复现 Dehejia-Wahba (1999) 里的 Table 1，看看每一个分组的 sample mean 。
均值和论文里的一致，standard error of sample mean 的数值和论文相似但是不完全一致，不知道论文中是用什么方式计算的。

```{r pre_treatment}
get.mean.se <- function(x) {
  mean = round(mean(x), 2)
  se = round(sd(x)/sqrt(length(x)), 2)
  return(paste0(mean, '(', se, ')'))
}
results <-  merge(
  data[, .(`no. obs`=.N), by = .(Dataset, treat)][Dataset == 'NSW' | treat ==0, ],
  data[, lapply(.SD, get.mean.se), by = .(Dataset, treat)][Dataset == 'NSW' | treat ==0, ]
)
kable(results[order(Dataset, -treat)])# caption = 'Sample Means of Characteristics for NSW and Comparison Samples')
```

## Regression 

目标是复现 Mostly Harmless Econometrics: Table 4.3.3 (page 68) 。

### Raw Difference

书中的 standard error 用的是 pooled variance。计算SE的时候学习了 Comparing Two Population Means: Independent Samples: 

- https://newonlinecourses.science.psu.edu/stat500/node/50/
- https://newonlinecourses.science.psu.edu/stat800/node/52/

```{r raw_difference}
raw.difference <- function(dt) {
  v1 <- dt[treat==1, ]$re78
  v0 <- dt[treat==0, ]$re78
  diff <- mean(v1) - mean(v0)
  n1 <- length(v1)
  n0 <- length(v0)
  sd1 <- sd(v1)
  sd0 <- sd(v0)
  var.pooled <- (n1 * sd1^2 + n0 * sd0^2) / (n1 + n0)
  se <- sqrt(var.pooled*(1/n1 + 1/n0))
  return(list('ATT'= round(diff, 0), 'SE' = round(se, 0), 'No. Obs.'= paste0(n1, '/', n0)))
}

results.rawdiff <- data[, raw.difference(.SD), by = .(Dataset)]
kable(results.rawdiff, escape = FALSE, caption = 'Raw Difference') 
```

### Demographic controls

这里复现的结果和书中数值不完全一致，但是趋势是差不多的。书里没有明确描述 regression 的时候具体的模型是如何的，例如有没有二次项。接下来几块利用了 demographic controls的结果都和书中数值不完全一致。

```{r demographic_controls}
# P-score Screened Samples
add.pscore.screened.samples <- function(fml) {
  cps1_pscore <- glm(fml, data = cps1, family=binomial(logit))
  cps1_pscore_screened <- cps1[(cps1_pscore$fitted.values >= 0.1 & cps1_pscore$fitted.values <= 0.9), ]
  cps3_pscore <- glm(fml, data = cps3, family=binomial(logit))
  cps3_pscore_screened <- cps3[(cps3_pscore$fitted.values >= 0.1 & cps3_pscore$fitted.values <= 0.9), ]
  cps1_pscore_screened[, Dataset := 'CPS-1-Subset']
  cps3_pscore_screened[, Dataset := 'CPS-3-Subset']
  return (rbind(data, cps1_pscore_screened, cps3_pscore_screened))
}

# Estimate the traning effects
estimate.via.regression <- function(dt, fml) {
  fit_summary <- summary(lm(fml, data = dt))
  diff <- coef(fit_summary)[2, 1]
  se <- coef(fit_summary)[2, 2]
  return(list('ATT'= round(diff, 0), 'SE' = round(se, 0), 'No. Obs.'= paste0(dt[treat==1, .N], '/', dt[treat==0, .N])))
}

fml_att <- re78 ~ treat + age + educ + black + hispan + nodegree + married
fml_pscore <- treat ~ age + educ + black + hispan + nodegree + married

data.tmp <- add.pscore.screened.samples(fml_pscore)
results.demographic <- data.tmp[, estimate.via.regression(.SD, fml_att), by = .(Dataset)]
kable(results.demographic, escape = FALSE, caption = 'Demographic controls') 
```

### 1975 Earnings

```{r 75_earnings}
fml_att <- re78 ~ treat + re75
fml_pscore <- treat ~ re75

data.tmp <- add.pscore.screened.samples(fml_pscore)
results.re75 <- data.tmp[, estimate.via.regression(.SD, fml_att), by = .(Dataset)]
kable(results.re75, escape = FALSE, caption = '1975 Earnings') 
```

### Demographics, 1975 Earnings

```{r demographic_controls_74}
fml_att <- re78 ~ treat + age + educ + black + hispan + nodegree + married + re75
fml_pscore <- treat ~ age + educ + black + hispan + nodegree + married + re75

data.tmp <- add.pscore.screened.samples(fml_pscore)
results.demographic.re75 <- data.tmp[, estimate.via.regression(.SD, fml_att), by = .(Dataset)]
kable(results.demographic.re75, escape = FALSE, caption = 'Demographics, 1975 Earnings') 
```

### Demographics, 1974 and 1974 Earnings

```{r demographic_controls_74_75}
fml_att <- re78 ~ treat + age + educ + black + hispan + nodegree + married + re74 + re75
fml_pscore <- treat ~ age + educ + black + hispan + nodegree + married + re74 + re75

data.tmp <- add.pscore.screened.samples(fml_pscore)
results.demographic.re7475 <- data.tmp[, estimate.via.regression(.SD, fml_att), by = .(Dataset)]
kable(results.demographic.re7475, escape = FALSE, caption = 'Demographics, 1974 and 1974 Earnings') 
```

### 汇总

```{r}
results.all <- results.demographic[, .(Dataset)]
results.all <- merge(results.all, results.rawdiff, by = 'Dataset', all.x = T, suffixes = c('', 'rawdiff'))
results.all <- merge(results.all, results.demographic, by = 'Dataset', all.x = T, suffixes = c('', 'demo'))
results.all <- merge(results.all, results.re75, by = 'Dataset', all.x = T, suffixes = c('', 're75'))
results.all <- merge(results.all, results.demographic.re75, by = 'Dataset', all.x = T, suffixes = c('', 'demo_re75'))
results.all <- merge(results.all, results.demographic.re7475, by = 'Dataset', all.x = T, suffixes = c('', 'demo_re7475'))
results.all <- results.all[, lapply(.SD, as.character)]
results.all[, 2:4][is.na(results.all[, 2:4])] <- ''
results.all[, 5:ncol(results.all)][is.na(results.all[, 5:ncol(results.all)])] <- 'no obs.'
kable(results.all[, c(1,seq(2, 16, 3))],
      escape = FALSE,
      caption = 'Regression estimates of NSW training effects using alternate controls',
      col.names = c('Dataset', 'Raw Difference',
                    'Demographic controls', '1975 Earnings', 'Demographics, 1975 Earnings', 'Demographics, 1974 and 1974 Earnings'))
```
